---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
library("badger")
knitr::opts_chunk$set(
  tidy= TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ShrinkCovMat: Shrinkage Covariance Matrix Estimators


[![Project Status: Active  The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![Travis-CI Build Status](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat.svg?branch=master)](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat)
`r badge_devel("AnestisTouloumis/ShrinkCovMat", "brightgreen")`
[![CRAN Version](http://www.r-pkg.org/badges/version/ShrinkCovMat?color=blue)](https://cran.r-project.org/package=ShrinkCovMat)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat?color=blue)](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat)
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-yellowgreen.svg)](/commits/master)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-2.11-6666ff.svg)](https://cran.r-project.org/)

## Author

Anestis Touloumis 

School of Computing, Engineering and Mathematics, University of Brighton 


## Installation

You can install the release version of the **ShrinkCovMat** package from CRAN:

```{r eval=FALSE}
install.packages("ShrinkCovMat")
```

of the development version from github:

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("AnestisTouloumis/ShrinkCovMat")
```

The source code for the **ShrinkCovMat** package is available on github at

- https://github.com/AnestisTouloumis/ShrinkCovMat.

To use **ShrinkCovMat**, you should load the package as follows:

```{r}
library(ShrinkCovMat)
```

## Usage and functions

The **ShrinkCovMat** package provides nonparametric Stein-type shrinkage estimators of the covariance matrix that are suitable and statistically efficient when the number of variables is larger than the sample size. These estimators are non-singular and well-conditioned regardless of the dimensionality.

Each of the implemented shrinkage covariance matrix estimators is a convex linear combination of the sample covariance matrix and of a target matrix. Three options are considered for the target matrix: 

- the identity matrix (`shrinkcovmat.identity`)
- the scaled identity matrix (`shrinkcovmat.equal`)
- the diagonal matrix with diagonal elements the corresponding sample variances (`shrinkcovmat.unequal`)

The utility function 

- `targetselection`

is designed to ease the selection of the target matrix.

## Example

The following R code illustrates how to use the core functions of **ShrinkCovMat**.

```{r}
data(colon)
normal.group <- colon[, 1:40]

Sigmahat1 <- shrinkcovmat.equal(normal.group)
Sigmahat1

Sigmahat2 <- shrinkcovmat.identity(normal.group)
Sigmahat2

Sigmahat3 <- shrinkcovmat.unequal(normal.group)
Sigmahat3
```

## How to cite

```{r}
citation("ShrinkCovMat")
```

## Contact
Want to help or have questions?  
Contact me directly, use an [issue](https://github.com/AnestisTouloumis/ShrinkCovMat/issues), fork me or submit a pull request.

[![star this repo](http://githubbadges.com/star.svg?user=AnestisTouloumis&repo=ShrinkCovMat&style=flat)](https://github.com/AnestisTouloumis/ShrinkCovMat)
[![fork this repo](http://githubbadges.com/fork.svg?user=AnestisTouloumis&repo=ShrinkCovMat&style=flat)](https://github.com/AnestisTouloumis/ShrinkCovMat/fork)