---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
library("badger")
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ShrinkCovMat: Estimation of a high-dimensional covariance matrix in R

[![Travis-CI Build Status](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat.svg?branch=master)](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat)
`r badge_devel("AnestisTouloumis/ShrinkCovMat", "brightgreen")`
[![CRAN Version](http://www.r-pkg.org/badges/version/ShrinkCovMat?color=blue)](https://cran.r-project.org/package=ShrinkCovMat)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat?color=blue)](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat)

## Author

Anestis Touloumis 

School of Computing, Engineering and Mathematics, University of Brighton 


## Installation

You can install the release version of the **ShrinkCovMat** package from CRAN:

```{r eval=FALSE}
install.packages("ShrinkCovMat")
```

of the development version from github:

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("AnestisTouloumis/ShrinkCovMat")
```

The source code for the **ShrinkCovMat** package is available on github at

- https://github.com/AnestisTouloumis/ShrinkCovMat.

## Functions

The **ShrinkCovMat** package provides three core function to estimate a covariance matrix.

- `shrinkcovmat.equal` `r badge_devel("AnestisTouloumis/ShrinkCovMat", "blue")`
- `shrinkcovmat.identity`
- `shrinkcovmat.equal`

The utility function 

- `targetselection`

helps users to choose the target matrix.

To use these functions, first, you should load the package as follows.

```{r}
library(ShrinkCovMat)
```
