---
output:
  md_document:
    variant: markdown_github
references:
- id: Touloumis2015
  title: Nonparametric Stein-type Shrinkage Covariance Matrix Estimators in High-Dimensional Settings
  author:
  - family: Touloumis
    given: Anestis
  container-title: Computational Statistics \& Data Analysis
  volume: 83
  URL: 'https://doi.org/10.1016/j.csda.2014.10.018'
  DOI: 10.1016/j.csda.2014.10.018
  page: 251-261
  type: article-journal
  issued:
    year: 2015
    month: March
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
library("badger")
knitr::opts_chunk$set(
  tidy= TRUE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# ShrinkCovMat: Shrinkage Covariance Matrix Estimators


[![Travis-CI Build Status](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat.svg?branch=master)](https://travis-ci.org/AnestisTouloumis/ShrinkCovMat)
`r badge_devel("AnestisTouloumis/ShrinkCovMat", "brightgreen")`
[![Project Status: Active The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active) 
[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](/commits/master)


[![CRAN Version](http://www.r-pkg.org/badges/version/ShrinkCovMat?color=blue)](https://cran.r-project.org/package=ShrinkCovMat)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat?color=blue)](http://cranlogs.r-pkg.org/badges/grand-total/ShrinkCovMat)
[![CRAN Downloads](http://cranlogs.r-pkg.org/badges/ShrinkCovMat)](http://cran.rstudio.com/web/packages/ShrinkCovMat/index.html)


## Author

Anestis Touloumis: <https://brighton.academia.edu/AnestisTouloumis>

School of Computing, Engineering and Mathematics, University of Brighton. 


## Installation

You can install the release version of the **ShrinkCovMat** package:

```{r eval=FALSE}
install.packages("ShrinkCovMat")
```

The source code of the release version is available on CRAN at:

- https://CRAN.R-project.org/package=ShrinkCovMat

Or you can install the development version of the **ShrinkCovMat** package:

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("AnestisTouloumis/ShrinkCovMat")
```

The source code for the **ShrinkCovMat** package is available on github at

- https://github.com/AnestisTouloumis/ShrinkCovMat.

To use **ShrinkCovMat**, you should load the package as follows:

```{r}
library(ShrinkCovMat)
```

## Usage

The **ShrinkCovMat** package provides nonparametric Stein-type shrinkage estimators of the covariance matrix that are suitable and statistically efficient when the number of variables is larger than the sample size. These estimators are non-singular and well-conditioned regardless of the dimensionality.

Each of the implemented shrinkage covariance matrix estimators is a convex linear combination of the sample covariance matrix and of a target matrix. Three options are considered for the target matrix: 

- the identity matrix (`shrinkcovmat.identity`),
- the scaled identity matrix (`shrinkcovmat.equal`),
- the diagonal matrix with diagonal elements the corresponding sample variances (`shrinkcovmat.unequal`).

Estimation of the corresponding optimal shrinkage intensities is discussed in [@Touloumis2015]. The utility function 

- `targetselection`

is designed to ease the selection of the target matrix.

## Example

The following R code illustrates how to use the core functions of **ShrinkCovMat**.

```{r}
data(colon)
normal.group <- colon[, 1:40]

Sigmahat1 <- shrinkcovmat.equal(normal.group)
Sigmahat1

Sigmahat2 <- shrinkcovmat.identity(normal.group)
Sigmahat2

Sigmahat3 <- shrinkcovmat.unequal(normal.group)
Sigmahat3
```

## How to cite

```{r}
citation("ShrinkCovMat")
```

# References